### 2016-07-13 appendix.tex
HTTP ongoing work

\section{Advanced Setup: HTTP Image Fetching}
\label{sec:http}

@@

\section{Alternative Commands for Outdated Packages}
This section describes alternative commands in case the testbed administrator
is not able to update old packages in its Linux distribution for performing
the commands, particularly the ones regarding the \texttt{util-linux} package.

@@

@@

>>>>

\section{HTTP Image Fetch Setup (Optional)}
\label{sec:http}

@@

\section{Alternative Commands for Outdated Packages}
This section describes alternative commands in case the testbed administrator
is not able to update old packages in its Linux distribution for performing
the commands, particularly the ones regarding the \texttt{util-linux} package
or if a command just fails.

@@

\subsection{Image Chroot with Proot}
\label{sec:chroot}

In some cases, \texttt{chroot} may not work properly. In this case, an
alternative can be \texttt{proot}. It might be required to install it
before changing the root with \texttt{apt-get install proot}. Then, it
should be possible for the testbed administrator to run the following:

% CHROOT to OS image
\begin{lstlisting}[]
# proot -q qemu-arm-static -S ${ROOTDIR}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

\subsection{Paramiko Package Update}
\label{sec:paramiko}
In case Paramiko version 1.15.1 or higher is not installed, we may observe
a key exchange algorithm error. In this case, it might be
necessary to update the Paramiko package and verify this by running:

\begin{lstlisting}[]
# pip install --upgrade paramiko
...
Sucesfully installed cffi-1.7.0 cryptography-1.4 paramiko-2.0.1
# pip show paramiko
...
Name: paramiko
Version: 2.0.1
...
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

This should allow to make \ac{SSH} connections with Fabric by using the proper
key exchange algorithms through Open SSH.

### 2016-07-13 conclusions.tex
HTTP ongoing work

\label{sec:conclusions}
Observing the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enables reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of this system. The description shows how to set up interconnected \ac{Raspi}s with small SD cards for local storage, a Raspberry distro image, network connectivity and proper system administration privileges. The exposed work will permit researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way. Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.

>>>>

\label{sec:conclusions}
Observing the expectation of the \ac{IoT} and lack for a low-cost, easy to configure testbed that enables reproducible results for mechanisms to address problems in this area, we provide a in-depth description of the new Aalborg University's \ac{Raspi} testbed for network coding evaluation and how to guarantee replicability and scaling managing of this system. The description shows how to set up interconnected \ac{Raspi}s with SD memory cards for local storage, a Raspbian Lite image, network connectivity and proper system administration privileges. We hope this work permits researchers to replicate setups and scenarios for evaluating their strategies in a rapid and manageable way. Future work in the use of \ac{Raspi} devices will focus on expanding the setup and automation of tasks to run the testbed, configure specified network topologies (e.g., with specific connectivity or packet loss ratios), reserve the use of these sub-networks for running tailored experiments and opening the use of the testbed beyond our team at Aalborg University.

### 2016-07-13 cross_compilation.tex
HTTP ongoing work

To use Kodo for research, it is required to obtain a research free license.
To do this, only a request form needs to be filled
in~\cite{steinwurflicenselink} and wait for grant access. Once
the access for Kodo has been granted, the source code can be pulled from
its Git repositories to be compiled. Thus, assuming that testbed
administrator already has access, we clone the Kodo repository locally
in \texttt{\${CODEDIR}} and change directory into the repository by doing:

>>>>

To use Kodo for research, it is required to obtain a research free license.
To do this, only a request form needs to be filled
in~\cite{steinwurflicenselink} and wait for grant access. Once
the access for Kodo has been granted, the source code can be pulled from
its Git repositories to be compiled. Thus, assuming that testbed
administrator already has access, we clone the \texttt{kodo-cpp}
repository locally in \texttt{\${CODEDIR}} and change directory into
the repository by doing:

### 2016-07-13 image_setup.tex
HTTP ongoing work

The output provides relevant information about the image. The image is in
total 2709504 sectors (1.3GiB) in size and contains two partitions. The
first partition starts at sector 8192 and the other partition starts at
sector 137216. The first partition type is FAT32 with a size of 63 MB
and the second partition is a Linux one with a size of 1.2 GB. This
indicates that the first partition is a boot partition and the second
one is a traditional Linux root filesystem, i.e. \texttt{/}.

\subsection{Image Resizing}
Given that we want to customize the files stored in the \ac{Raspi}s,
we need to resize the image file since 1.2GB might not be enough to store
the root filesystem due to the total size of the additional packages. Thus,
we need to increase the partition size. The following procedure illustrates
how the image and its root filesystem can be expanded by one \ac{GB}.
First, to expand the image one \ac{GB}, we execute:

@@

Later, we use \texttt{fdisk} with the same arguments as before to see that
the image is now one \ac{GB} larger:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

Now, in this command block output, we observe that the change has taken
effect by noticing the total available image size is 2.3 GiB. To expand the
root filesystem, it is required to first remove the Linux partition
and then add it again with one \ac{GB} more. To do this, we make use
again of \texttt{fdisk} in command mode to alter the partition table.
For this, we pass the arguments of the command mode of \texttt{fdisk}
through the \texttt{echo} command in Linux and the \texttt{|} operator
(pipe) as follows:

@@

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
-----------------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

%\url{https://wiki.debian.org/QemuUserEmulation}
%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
\subsection{Image Package Updating}
Besides adding and configuring files within the image, the testbed
administrator may want to install and update the software packages
within the image before it is written to all the memory cards that goes
into the \ac{Raspi}s. From any Linux x86 machine as the testbed administrator
\ac{PC}, this can be done using \texttt{chroot} command in the
QEMU~\cite{QemuUserEmulation} hypervisor for \ac{ARM} processors.

@@

% In some cases, \texttt{chroot} may not work properly. tgus, in this case, an
% alternative can be \texttt{proot}. It may be required to install it
% before changing the root with \texttt{apt-get install proot}. Then, it
% should possible for the testbed administrator to run the following:

% % CHROOT to OS image
% \begin{lstlisting}[]
% # proot -q qemu-arm-static -S ${ROOTDIR}
% \end{lstlisting}
% \FloatBarrier
% \vspace{-5mm}
If successfully executed, our terminal should have changed the prompt
indicating that we are the root user in the Raspbian lite root filesystem as
the apparent root. To be aware of the mode that we are working now, we change
the prompt title to indicate that it is a \texttt{chroot} environment as
follows:

>>>>

The output provides relevant information about the image. The image is in
total 2709504 sectors (1.3 GiB) in size and contains two partitions. The
first partition starts at sector 8192 and the other partition starts at
sector 137216. The first partition type is FAT32 with a size of 63 MB
and the second partition is a Linux one with a size of 1.2 GB. This
indicates that the first partition is a boot partition and the second
one is a traditional Linux root filesystem, i.e. \texttt{/}.

\subsection{Image Resizing}
Given that we want to customize the files stored in the \ac{Raspi}s,
we need to resize the image file since 1.2 GB might not be enough to store
the root filesystem due to the total size of the additional packages. Thus,
we need to increase the partition size. The following procedure illustrates
how the image and its root filesystem can be expanded by one GB.
First, to expand the image one GB, we execute:

@@

Later, we use \texttt{fdisk} with the same arguments as before to see that
the image is now one GB larger:
\begin{lstlisting}[literate={DOLLAR}{\$}1]
DOLLAR fdisk -u sectors -l ${IMAGE}.img
Disk 2016-05-27-raspbian-jessie-lite.img: 2.3 GiB, 2461007872 bytes, 4806656 sectors
Units: sectors of 1 * 512 = 512 bytes
Sector size (logical/physical): 512 bytes / 512 bytes
I/O size (minimum/optimal): 512 bytes / 512 bytes
Disklabel type: dos
Disk identifier: 0x6fcf21f3

@@

Now, in this command block output, we observe that the change has taken
effect by noticing the total available image size is 2.3 GiB. To expand the
root filesystem, it is required to first remove the Linux partition
and then add it again with one GB more. To do this, we make use
again of \texttt{fdisk} in command mode to alter the partition table.
For this, we pass the arguments of the command mode of \texttt{fdisk}
through the \texttt{echo} command in Linux and the \texttt{|} operator
(pipe) as follows:

@@

\Suppressnumber\begin{lstlisting}[]
<@\textcolor{gray}{\$\{ROOTDIR\}/etc/rc.local}@>
<@\textcolor{gray}{
---------------------------------------------}
\Reactivatenumber @>
...
bash /home/pi/rasp_config/update_rasp_config
bash /home/pi/rasp_config/main
...
exit 0
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

%\url{https://wiki.debian.org/QemuUserEmulation}
%\url{https://wiki.archlinux.org/index.php/Raspberry_Pi}
\subsection{Image Packages Updating by Changing the Apparent Root Directory}
Besides adding and configuring files within the image, the testbed
administrator may want to install and update the software packages
within the image before it is written to all the memory cards that goes
into the \ac{Raspi}s. From any Linux x86 machine as the testbed administrator
\ac{PC}, this can be done using \texttt{chroot} command in the
QEMU~\cite{QemuUserEmulation} hypervisor for \ac{ARM} processors.

@@

If successfully executed, our terminal should have changed the prompt
indicating that we are the root user in the Raspbian lite root filesystem as
the apparent root. In case the \texttt{chroot} command is not successful,
we provide an alternative command in Section~\ref{sec:chroot} of the Appendix.
To be aware of the mode that we are working now, we change
the prompt title to indicate that it is a \texttt{chroot} environment as
follows:

### 2016-07-13 testbed_http.tex
Changing Chapter 4 structure
HTTP ongoing work

\label{sec:testbed_http}

\subsection{HTTP}

%This part will explain how to prepare the existing image to be downloaded from
%a \ac{HTTP} server.
The Linux image that has been made so far has two partitions. One 
partition with the boot files and one containing the root file system.

This section will present how the partitions can be extracted from the image,
and how the boot partition can be configured to download the root filesystem
from a \ac{HTTP} server during startup.

The advantage of this setup is that multiple \ac{Raspi}s can share the same
root image. This significantly eases the maintainance task.

%An alternative to the setup explained in this procedure is to use \ac{NFS}
%instead of \ac{HTTP}.

%The first program called when a \ac{Raspi} starts is the Boot Loader,
%\texttt{U-Boot}. \texttt{U-Boot} 
%
%When Raspbian starts up, the first program called
%
%The procedure explained below will customize the initial \ac{RAM} filesystem
%
%
%there exists only one root file
%
%The advantage of this is that the root filesystem only one all \ac{Raspi}s can automatically download the 
%
%The boot partition will be required in the \ac{Raspi}s, but the root image can

Start by declaring the path and names of the boot- and root-image:. 
\begin{lstlisting}[]
$ export HTTPBOOTIMAGE="${WORKDIR}/httpboot"
$ export HTTPROOTIMAGE="${WORKDIR}/httproot"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The, setup a loopback device of the existing image and use \texttt{dd} to
extract the boot and root partitions from the image:
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
# dd if=${DEV}p1 of=${HTTPBOOTIMAGE}.img bs=4M
# dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
# losetup -d ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

In a previous steps, the boot loader was configured to call an initial \ac{RAM}
filesystem. The coming steps will present how to the initial \ac{RAM}
filesystem can be extended further to download and cache the root filesystem
from a \ac{HTTP} server before mounting it.

>>>>

\label{sec:testbed_http}
Creating manual copies of the customized image for each memory card can
be also a time-consuming task that is prone to errors and become
cumbersome. Therefore, we present an optional more advanced \ac{HTTP}
image fetching setup for the existing image to be downloaded from a
\ac{HTTP} server. We consider this an optional step since a testbed
administrator might opt for not including this as part of his / her setup.
We start our description consideri

At the end of Section~\ref{sec:overlay_fs}, the final customized image was
proposed to be splitted into three partitions. One for the boot files, one
containing the root file system and one for the home directory. This is made
considering an overlay filesystem approach. In this section, instead of
writing the image direclty to the memory cards, we present how the
partitions can be extracted from the image, and how the boot partition can
be configured to download the root filesystem and home directory partitions
from a \ac{HTTP} server during startup. The advantage of this setup is that
we completely automate the image fetch which significantly eases its
maintainance since any changes are required to be made only in the \ac{HTTP}
server image. We start by declaring the variables for the path names of each
partition:

\begin{lstlisting}[]
$ export HTTPBOOTIMAGE="${WORKDIR}/httpboot"
$ export HTTPROOTIMAGE="${WORKDIR}/httproot"
$ export HTTPHOMEIMAGE="${WORKDIR}/httphome"
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Then, we setup a loopback device of the existing image and use \texttt{dd} to
extract the boot and root partitions from the image:
\begin{lstlisting}[]
# export DEV=$(sudo losetup --show -f -P ${IMAGE}.img); echo $DEV
# dd if=${DEV}p1 of=${HTTPBOOTIMAGE}.img bs=4M
# dd if=${DEV}p2 of=${HTTPROOTIMAGE}.img bs=4M
# losetup -d ${DEV}
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

@@

In a previous steps, the boot loader was configured to call an initial \ac{RAM}
filesystem. Now, we present how to the initial \ac{RAM}
filesystem can be extended further to download and cache the root filesystem
from a \ac{HTTP} server before mounting it.

### 2016-07-13 tools.tex
HTTP ongoing work

The previous \texttt{fabfile} shows three function that perform our
example tasks. These functions utilize variables and subsequent functions
from the Fabric \ac{API} such as \texttt{env}, \texttt{task} and \texttt{sudo}
among others. Each of these \ac{API} functions permit to define environment
variables, create the administrator tasks through decorators or run the
mentioned task in \texttt{sudo} mode, respectively.
When a task is called from the terminal, Fabric searches the directory for
the \texttt{fabfile.py} file and executes the desired task. The syntax for
executing a task with arguments is in the form
\texttt{fab <TASK>:arg1,arg2,...}. Some executions from the terminal
are shown as follows \todot{Test the following commands later and print the
outputs}:

\begin{lstlisting}[]
$ fab reboot
$ fab install:tmux
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The first function above reboots the \ac{Raspi}s in the lists of hosts and
the second function installs a program given by an argument. In both cases,
after a successful \ac{SSH} connection is made, Fabric employs the
\ac{Raspi}'s \texttt{reboot} and \texttt{apt-get} commands in \texttt{sudo}
mode to do the required tasks. Below it is shown an example for
the \texttt{push} task which uses two arguments. Here we copy
\texttt{my\_file} from the testbed administrator \ac{PC} to all the host
\ac{Raspi}s:

\begin{lstlisting}[]
$ fab push:"${CODEDIR}/my_file",'~/'
[192.168.87.107] Executing task 'push'
[192.168.87.107] put: /home/<USER>/code/my_file -> /home/pi/my_file

Done.
Disconnecting from 192.168.87.107... done.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

Fabric has many other functionalities that are useful in
controlling a large set of \ac{Raspi}s. For example, we may extract files
or run automated experiments. The included functionalities in
the \texttt{\$CODEDIR/fabfile.py} file will depend on the requirements
of the testbed administrator.

@@

There are methods to enable the \ac{Raspi}s to continue running
applications although the connection is terminated either on purpose or
unexpectedly. One method is to run programs within a \texttt{screen} session.
\texttt{screen} enables a user to run applications within a shell
window, a \texttt{screen} session, which does not terminate even with
connectivity interruptions. Users can attach and detach from such a session
as desired. The following procedure presents how to use \texttt{screen} with
\ac{SSH} to: (i) Login to a generic \ac{Raspi} with \ac{IP} address as
\texttt{<RASP\_IP>}, (ii) open a screen session, (iii)
execute an example command, (iv) detach from the screen session, (v)
terminate the \ac{SSH} connection, (vi) login to the \ac{Raspi} again
and (vii) attach to screen session to see the program still running.
From the testbed administrator, we start by establishing \ac{SSH} connection
to a \ac{Raspi} and open a screen session:

>>>>

The previous \texttt{fabfile} shows three function that perform our
example tasks. These functions utilize variables and subsequent functions
from the Fabric \ac{API} such as \texttt{env}, \texttt{task} and \texttt{sudo}
among others. Each of these \ac{API} functions permit to define environment
variables, create the administrator tasks through decorators or run the
mentioned task in \texttt{sudo} mode, respectively.
When a task is called from the terminal, Fabric searches the directory for
the \texttt{fabfile.py} file and executes the desired task. The syntax for
executing a task with arguments is in the form
\texttt{fab <TASK>:arg1,arg2,...}. In what follows, we denote the
\ac{IP} address of a generic \ac{Raspi} for test as \texttt{<RASP\_IP>}.
The executions from the terminal of some of these commands are shown as
follows:

\begin{lstlisting}[]
$ fab reboot
[<RASP_IP>] Executing task 'reboot'

Done.
Disconnecting from <RASP_IP>... done.
$ fab install:tmux
[<RASP_IP>] Executing task 'install'
...
The following NEW packages will be installed:
  tmux
0 upgraded, 1 newly installed, 0 to remove and 0 not upgraded.
...
Preparing to unpack .../archives/tmux_1.9-6_armhf.deb ...
Unpacking tmux (1.9-6) ...
Processing triggers for man-db (2.7.0.2-5) ...
Setting up tmux (1.9-6) ...

Done.
Disconnecting from <RASP_IP>... done.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

The first function above reboots the \ac{Raspi}s in the lists of hosts and
the second function installs a program given by an argument. For the
connection to the devices, Fabric calls the Paramiko module from Python
to make a \ac{SSH} connection. For this to work properly, the Paramiko
version needs to be higher than or equal to 1.15.1. If not available,
the \ac{SSH} connections from Fabric may fail. Thus in case of any problems,
some instructions for updating the Paramiko package are available in
Section~\ref{sec:paramiko} of the Appendix. This is a standard recommendation
the Fabric troubleshooting guide~\cite{2016fabricsupport}.

After a successful \ac{SSH}
connection is made, in the previous two commands towards the \ac{Raspi},
Fabric employs the \ac{Raspi}'s \texttt{reboot} and \texttt{apt-get}
commands in \texttt{sudo} mode to do the required tasks. Below it is shown
an example for the \texttt{push} task which uses two arguments. Here we
copy \texttt{my\_file} from the testbed administrator \ac{PC} to a test
host \ac{Raspi}:

\begin{lstlisting}[]
$ fab push:"${CODEDIR}/my_file",'~/'
[<RASP_IP>] Executing task 'push'
[<RASP_IP>] put: /home/<USER>/code/my_file -> /home/pi/my_file

Done.
Disconnecting from <RASP_IP>... done.
\end{lstlisting}
\FloatBarrier
\vspace{-5mm}

To control a large set of devices, we simply need to include them in the
\texttt{env.hosts} list in the \texttt{\$CODEDIR/fabfile.py} file. Fabric has
many other functionalities that are useful in controlling a large set of
\ac{Raspi}s. For example, we may extract files or run automated
experiments. The included functionalities in the
\texttt{\$CODEDIR/fabfile.py} file will depend on the requirements of the
testbed administrator.

@@

There are methods to enable the \ac{Raspi}s to continue running
applications although the connection is terminated either on purpose or
unexpectedly. One method is to run programs within a \texttt{screen} session.
\texttt{screen} enables a user to run applications within a shell
window, a \texttt{screen} session, which does not terminate even with
connectivity interruptions. Users can attach and detach from such a session
as desired. The following procedure presents how to use \texttt{screen} with
\ac{SSH} to: (i) Login to a generic \ac{Raspi}, (ii) open a screen session,
(iii) execute an example command, (iv) detach from the screen session, (v)
terminate the \ac{SSH} connection, (vi) login to the \ac{Raspi} again
and (vii) attach to screen session to see the program still running.
From the testbed administrator, we start by establishing \ac{SSH} connection
to a \ac{Raspi} and open a screen session:
